---
title: "Data Wrangling I"
output: github_document
---
# Data Import
Load necessary packages for this file.
```{r load packages, message = FALSE}
library(tidyverse)
```

Let's import the `FAS_litters.csv` csv ysubg a relative path.
```{r data import}
#Load in using read_csv
litters_df =
  read_csv(file = "data/FAS_litters.csv")

# Clean column names, note "::" accesses a specific function within an unloaded package
litters_df = 
  janitor::clean_names(litters_df)

```

Let's import the `FAS_litters.csv` csv using an absolute path.

```{r data import absolute, eval = FALSE}
#Load in using read_csv
litters_df_absolute =
  read_csv(file = "~/Documents/Data Science/data_wrangling_i/data/FAS_litters.csv")

# Clean column names
litters_df_absolute = 
  janitor::clean_names(litters_df_absolute)

```

Don't use absolute paths. It makes your life more complicated when you try to move/rename/distribute your work.

### Learning Assessment

I will load in info about the pups using relative pathing.
```{r load pups_df}
pups_df = 
  read_csv("data/FAS_pups.csv")

# Clean column names
pups_df = 
  janitor::clean_names(pups_df)

pups_df
```

## Look at data

```{r}
litters_df

head(litters_df)

tail(litters_df)
```

You can use `view`. But don't include in RMD files, use in command line/non-executed code.
```{r view, eval = FALSE}
# Pops open another window, weird interactions with 'knit'
view(litters_df)
```


Look at data structure.
```{r}
str(litters_df)

skimr::skim(litters_df)
```


